A Case Study in Preserving a\\High Energy Physics Application

### 2014-03-21 paper.tex
merge Dr. Hildreth's comments

\begin{abstract}
\it The reproducibility of scientific results increasingly
depends upon the preservation of computational artifacts.
Although preserving a computation to be used later sounds
easy, it is surprisingly difficult due to the complexity
of existing software and systems.  Implicit dependencies,
networked resources, and shifting compatibility all conspire
to break applications that appear to work well.  To investigate
these issues, we present a case study of a complex high energy
physics application.  We analyze the application and attempt
several methods at extracting its dependencies for the purposes
of preservation. 
We propose one fine-grained dependency management toolkit to preserve the application and demonstrate its correctness in two different environments - one virtual machine from Notre Dame Cloud Platform and one virtual machine from Amazon EC2 Platform. 
We report on the completeness, performance,
and efficiency of each technique, and offer some guidance for
future work in application preservation.
\end{abstract}

@@

In a very abstract sense, reproducing a computation is trivial.
Assuming a computation is deterministic, one must simply
preserve all of the inputs to a computation, then re-run
the same code in an equivalent environment, and the same result
will be produced.  For a small custom application on a modest
amount of data, this could be accomplished by capturing the
complete environment,
data, and software within a single virtual machine image,
and then depositing the virtual
it into a curated environment.  The publication could
then simply refer to the identifier of the image, which the
interested reader can obtain and re-use. This approach has
been used to some success with systems~\cite{castagne2013consider}.
\footnote{Of course, we are glossing over the problem that hardware
architectures and virtual machines also change, so one must also
preserve the VMM software necessary to run the image.  The VMM itself
depends on a software environment which must also be preserved.
A long-term preservation system might end up running a whole
stack of nested virtual machines in order to provide the desired
environment! }

However, this simple approach is not sufficient for large applications
that are run in complex social environments.

@@

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN~\cite{collaboration2008cms}, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks.  As the Higgs boson is to short-lived to be detected
itself, it has to be reconstructed from its decay products.

The application which is the study of this paper is called \emph{TauRoast}.
It searches for cases where the Higgs boson decays to two tau leptons.
The leptons are not observed directly, but by the particle showers
that they generate.  So, the analysis must search for detector
events that show a signature of decay products compatible with both hadronic tau and top decays.  Properties of such events are used to distinguish
the events of interest (Higgs decays) from all other events and
are also used in further statistical analysis.

@@

The largest of these repositories is the CMS Software Distribution (CMSSW),
a carefully-curated selection of software packages which is distributed
in several forms.  Historically, components of CMSSW were obtained by checking components
of the source out of CVS, or by installing a complete binary package on a shared
filesystem within an HPC center.  In recent years, distribution has moved to
an on-demand delivery system known as CVMFS~\cite{blomer2011cernvm}, which
provides a filesystem interface that transparently accesses a remote repository.
The content of CMSSW is managed very carefully by a centralized team whose main goal
is to ensure that the current version of the software operates correctly
on the operating systems and architectures currently in use.  However,
preservation is not a specific objective of the system, and so there is
no particular guarantee that old versions of CMSSW will continue to
operate indefinitely.

CMSSW contains many different tools, libraries, and utilities.  No single code uses anywhere close to all of that.  But, because it is widely used within the community, it is common for users and developers to simply expect that a particular version of the entire repository is available.

{\bf Data Sources.}
The CMS collaboration provides analysis end-users with a pre-processed
and reduced data format, AOD~\cite{holtman2001cms}, containing information for events, i.e.,
proton-proton collisions with a signature of interest, in the form of
reconstructed particles.  This format is based on the RAW output of
the CMS detector readout electronics and reconstructed world-wide.
Both real and simulated data are available for examination.

As AOD data are too large to be iteratively processed repetitively in
an physics analysis workflow, it is normally reduced further in
structural complexity and content.  For the analysis under
investigation here, this is a two-step process.  First, the AOD data
are processed at the Notre Dame working group cluster to BEAN events,
containing only trivial data containers packed in vectors.  This step
is time and CPU intensive and its output contains data of 11.6$\,$TB to be
analyzed by the tau analysis.
It is performed by a small custom code framework, which is built on top of CMSSW.
The BEAN format, production code, and
data are shared within the analysis group looking at Higgs production
in association with top quarks, which is formed by groups from a few
American and European universities,
consisting of up to a few dozen contributors.

@@

First, we elaborated the email instructions into an
executable script that obtains the dependencies and then
executes the analysis.  The script declares the necessary
environment variables, downloads and checks out the necessary source code,
builds it appropriately, calls initialization scripts in
the dependent software, and then runs the analysis.
A few rounds of correction with the original author were necessary
to obtain all the dependencies and run the artifact correctly.
(The original email also indicated how to run the application
within a production batch system.  For the purposes of preservation,
we consider the execution infrastructure to be distinct from the application,
and leave it out of consideration for now.)

@@

\item {\bf Many Implicit Local Dependencies.} A much harder problem is that the
application assumed the presence of many different components in the local
filesystem view. It would be tempting to capture all of these by simply
storing a virtual machine image containing the local filesystem. However,
the application depended on no less than {\bf five} networked filesystems
available on a particular machine available to the author:
the data to be analyzed was stored on a HDFS~\cite{borthakur2008hdfs} cluster,
some configuration data was stored on a CVMFS~\cite{blomer2011cernvm} filesystem,
and a variety of software tools were on a NFS~\cite{howard1988scale},
PanFS~\cite{welch2008scalable} and AFS~\cite{sandberg1985design} systems.
The original authors were not aware of many of these dependencies,
because they simply relied on local administrators to configure the
software and make it available.

@@

\item {\bf High Selectivity.}  Although the total size of the resources accessed by this
program is very large, the size of the data and software actually used are much smaller.
Often, an entire repository or data source is named within the script, but the program
only needs a handful of items from that source.  For example, the data is stored on an
HDFS filesystem with 11.6TB of data, but only 20GB are actually consumed by the program. The CMSSW repository is 88.1GB in total
but only 448.3MB in source are checked out, and the software actually used only
measured 6.3MB.  In a few cases, a source of software is named but never actually accessed.
(For example, our original script includes the Open Science Grid software stack in the PATH, but does not actually use it.)
We suspect that end users are accustomed to missing dependencies and thus get in the habit of adding commonly used software,
whether it is needed or not.

@@

Figure~\ref{fig:workflow-parrot} illustrates the measurement process.
The starting point of this toolkit is one successful execution of the application on the native machine.
First, we execute the actual data analysis script under Parrot to generate the namelist.
Then, using the namelist, we generate a package containing all the necessary data
and software for one analysis program is generated. When another
researcher wants to repeat the program, he only needs to obtain the package and
execute the actual analysis program inside the package.

@@

We evaluated the correctness of the reduced package and the overhead of generating the package.
To do it, the application was repeated with the help of Original Script (as shown in Figure~\ref{fig:version-evolution}) on the original machine, mentioned as {\bf Original-Script} in the following part, 
one reduced package was generated and verified with the help of Parrot on the original machine, which will be mentioned as {\bf Reduced-Package}.
Then, two different virtual machines (VM) - one VM from Notre Dame Cloud Platform based on KVM and one VM from Amazon EC2 Platform based on Xen, were employed to further verify the correctness of the reduced package.

@@

To measure time consumption, we counted the time used to obtain remote software dependencies, build environment, and analyze the dataset respectively. We also counted the time used to obtain the filename list and generate the package. To measure data size, we can easily figure out the size of each data and code source inside the package under Reduced-Package.
Original-Script does not support mining of implicit dependencies. As for remote sources, we can figure out the named size through the analysis script. However, it is hard to figure out the named size of each local source. 
Instead, we only knew that total size of each file system is very large.

@@

Table~\ref{table:config-vm} illustrates the configuration of 
each machine and the execution time of the application on each machine.
All the machines adopt x86\_64 hardware platform and Linux OS.
Both of the two VMs repeated the application with the help of the package generated on the original machine successfully.
The execution time on one machine greatly depends on its hardware configuration.

@@

This work was supported in part by National Science Foundation grants PHY-1247316 (DASPOS)
and OCI-1148330 (SI2).  We also thank...

>>>>

\begin{abstract}
\it The reproducibility of scientific results increasingly
depends upon the preservation of computational artifacts.
Although preserving a computation to be used later sounds
easy, it is surprisingly difficult due to the complexity
of existing software and systems.  Implicit dependencies,
networked resources, and shifting compatibility all conspire
to break applications that appear to work well.  To investigate
these issues, we present a case study of a complex high energy
physics application.  We analyze the application and attempt
several methods at extracting its dependencies for the purposes
of preservation. 
We propose one fine-grained dependency management toolkit to preserve the application and demonstrate its correctness in two different environments - one virtual machine from the Notre Dame Cloud Platform and one virtual machine from the Amazon EC2 Platform. 
We report on the completeness, performance,
and efficiency of each technique, and offer some guidance for
future work in application preservation.
\end{abstract}

@@

In a very abstract sense, reproducing a computation is trivial.
Assuming a computation is deterministic, one must simply
preserve all of the inputs to a computation, then re-run
the same code in an equivalent environment, and the same result
will be produced.  For a small custom application on a modest
amount of data, this could be accomplished by capturing the
complete environment,
data, and software within a single virtual machine image,
and then depositing the virtual
image into a curated environment.  The publication could
then simply refer to the identifier of the image, which the
interested reader can obtain and re-use. This approach has
been used to some success with systems~\cite{castagne2013consider}.
\footnote{Of course, we are glossing over the problem that hardware
architectures and virtual machines also change, so one must also
preserve the VMM software necessary to run the image.  The VMM itself
depends on a software environment which must also be preserved.
A long-term preservation system might end up running a whole
stack of nested virtual machines in order to provide the desired
environment! }

However, this simple approach is not sufficient for large applications
that are run in complex environments.

@@

Within the ongoing investigation of the Higgs boson at the CMS
detector, part of the LHC at CERN~\cite{collaboration2008cms}, the Higgs production in association
with two top quarks allows to measure the Higgs coupling strength to
top quarks~\cite{chatrchyan2013search}.  As the Higgs boson is too short-lived to be detected
itself, it has to be reconstructed from its decay products.

The application which is the study of this paper is called \emph{TauRoast}.
It searches for cases where the Higgs boson decays to two tau leptons.
Since the tau leptons are very short-lived, they are not observed directly, but by the particle decay products 
that they generate.  So, the analysis must search for detector
events that show a signature of decay products compatible with both hadronic tau and top decays.  Properties of such events are used to distinguish
the events of interest (Higgs decays) from all other events and
are also used in further statistical analysis.

@@

The largest of these repositories is the CMS Software Distribution (CMSSW)~\cite{cms2006cms},
a carefully-curated selection of software packages which is distributed
in several forms.  Historically, components of CMSSW were obtained by checking components
of the source out of CVS (CMSSW source codes were moved from CVS to git recently), or by installing a complete binary package on a shared
filesystem within an HPC center.  In recent years, distribution has moved to
an on-demand delivery system known as CVMFS~\cite{blomer2011cernvm}, which
provides a filesystem interface that transparently accesses a remote repository.
The content of CMSSW is managed very carefully by a centralized team whose main goal
is to ensure that the current version of the software operates correctly
on the operating systems and architectures currently in use.  However,
preservation is not a specific objective of the system, and so there is
no particular guarantee that old versions of CMSSW will continue to
operate indefinitely.

CMSSW contains many different tools, libraries, and utilities.  No single code uses anywhere close to all of these.  But, because it is widely used within the experimental researchers, it is common for users and developers to simply expect that a particular version of the entire repository is available.

{\bf Data Sources.}
The CMS collaboration provides analysis end-users with a pre-processed
and reduced data format, AOD~\cite{holtman2001cms}, containing information for events, i.e.,
proton-proton collisions with a signature of interest, in the form of
reconstructed particles.  This format is based on the RAW output of
the CMS detector readout electronics and reconstructed world-wide, which is then processed through various algorithms derive signatures of individual particles.
Both real and simulated data are available for examination.

As AOD data are too large to be iteratively processed repetitively in
a physics analysis workflow, it is normally reduced further in
structural complexity and content.  For the analysis under
investigation here, this is a two-step process.  First, the AOD data
are processed at the Notre Dame working group cluster to BEAN (Boson Exploration Analysis Ntuple) events,
containing only trivial data containers packed in vectors.  This step
is time and CPU intensive and its output contains data of 11.6$\,$TB to be
analyzed by the tau analysis.
It is performed by a small custom code framework, which is built on top of CMSSW.
The BEAN format, production code, and
data are shared within the analysis group looking at Higgs production
in association with top quarks, which is formed by groups from a few
American and European universities,
consisting of up to a few dozen contributors.

@@

First, we elaborated the email instructions into an
executable script that obtains the dependencies and then
executes the analysis.  The script declares the necessary
environment variables, downloads and checks out the necessary source code,
builds it appropriately, calls initialization scripts in
the dependent software, and then runs the analysis.
A few rounds of correction with the original author were necessary
to obtain all the dependencies and run the artifact correctly.
(The original instructions introduced in the email also indicated how to run the application
within a production batch system.  For the purposes of preservation,
we consider the execution infrastructure to be distinct from the application,
and leave it out of consideration for now.)

@@

\item {\bf Many Implicit Local Dependencies.} A much harder problem is that the
application assumed the presence of many different components in the local
filesystem view. It would be tempting to capture all of these by simply
storing a virtual machine image containing the local filesystem. However,
the application depended on no less than {\bf five} networked filesystems
available on a particular machine available to the author:
the data to be analyzed was stored on an HDFS~\cite{borthakur2008hdfs} cluster,
some configuration data was stored on a CVMFS~\cite{blomer2011cernvm} filesystem,
and a variety of software tools were on an NFS~\cite{howard1988scale},
PanFS~\cite{welch2008scalable} and AFS~\cite{sandberg1985design} systems.
The original authors were not aware of many of these dependencies,
because they simply relied on local administrators to configure the
software and make it available.

@@

\item {\bf High Selectivity.}  Although the total size of the resources accessed by this
program is very large, the size of the data and software actually used are much smaller.
Often, an entire repository or data source is named within the script, but the program
only needs a handful of items from that source.  For example, the data is stored on an
HDFS filesystem with 11.6TB of data, but only 20GB are actually consumed by the program. 
The reason for this great reduction is at first each BEAN event contains a large amount of information, and the TauAnalysis throws away a lot of the irrelevant event information, keeping only the relevant bits.
The CMSSW repository is 88.1GB in total
but only 448.3MB in source are checked out, and the software actually used only
measured 6.3MB.  In a few cases, a source of software is named but never actually accessed.
(For example, our original script includes the Open Science Grid software stack in the PATH, but does not actually use it.)
We suspect that end users are accustomed to missing dependencies and thus get in the habit of adding commonly used software,
whether it is needed or not.

@@

Figure~\ref{fig:workflow-parrot} illustrates the measurement process.
The starting point of this toolkit is one successful execution of the application on the native machine.
First, we execute the actual data analysis script under Parrot to generate the namelist.
Then, using the namelist, we generate a package containing all the necessary data
and software for one analysis program. When another
researcher wants to repeat the program, he only needs to obtain the package and
execute the actual analysis program inside the package.

@@

We evaluated the correctness of the reduced package and the overhead of generating the package.
To do this, the application was repeated with the help of Original Script (as shown in Figure~\ref{fig:version-evolution}) on the original machine. 
Then one reduced package was generated and verified with the help of Parrot on the original machine.

Then, two different virtual machines (VM) - one VM from the Notre Dame Cloud Platform based on KVM and one VM from the Amazon EC2 Platform based on Xen, were employed to further verify the correctness of the reduced package.

@@

To measure time consumption, we counted the time used to obtain remote software dependencies, build the environment, and analyze the dataset. We also counted the time used to obtain the filename list and generate the package. To measure data size, we can easily figure out the size of each data and code source inside the package under Reduced-Package.
Original-Script does not support mining of implicit dependencies. As for remote sources, we can figure out the named size through the analysis script. However, it is hard to figure out the named size of each local source. 
Instead, we only knew that total size of each file system is very large.

@@

Table~\ref{table:config-vm} illustrates the configuration of 
each machine and the execution time of the application on each machine.
All the machines adopt the x86\_64 hardware platform and Linux OS.
Both of the two VMs repeated the application with the help of the package generated on the original machine successfully.
The execution time on one machine greatly depends on its hardware configuration.

@@

This work was supported in part by National Science Foundation grants PHY-1247316 (DASPOS), 
OCI-1148330 (SI2) and PHY-1312842.  We also thank the support and help of the Center for Research Computing (CRC) of University of Notre Dame.
